---
title: "STAT-180 Project-1"
author: "Sophia Zhang, Mojdeh Motalebi Kashani"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```

# Q2
First, we import that dataset and remove and filter-out any rows with missing variables.
```{r}
survey <- read.csv("survey.csv")
survey[survey == ''] <- NA
survey <- survey %>% drop_na()
```

The first contingency table compares the level of exercise and smoking in students. Majority of people never smoke. Also, even among those that smoke, most exercise sometimes or frequently.
```{r}
table(data.frame("Exer" = survey$Exer, "Smoke" = survey$Smoke))
```

The second contingency table compares writing hand with the upper hand when clapping. Most students are right-handed and most right-handed students clap with their right hand on top.
```{r}
table(data.frame("W.Hnd" = survey$W.Hnd, "Clap" = survey$Clap))
```

The first graph is a boxplot of student heights. With a median of 171, a max of 200, and a min of 150.  
```{r}
boxplot(survey$Height, main = "Student Height")
```

The second graph is a histogram of student ages. Most people are under 20, vast majorty of people are under 25, and the oldest people are over 70.
```{r}
hist(survey$Age, main = "Student Age")
```

Finally, you can see the summary statistics for 2 variables. Student pulse rate(beats per minute) and student writing hand span(distance from tip of thumb to tip of little finger of spread hand).
```{r}
summary(survey$Pulse)
summary(survey$Wr.Hnd)
```