---
title: "STAT-180 Project-1"
author: "Sophia Zhang, Mojdeh Motalebi Kashani"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(hrbrthemes)
library(geomtextpath)
```
First step, find the name of all the variables inside of the list. 
```{r setup, include=TRUE}
survey$rownames <- NULL ##This removes the first column which is just the number of rows. 
names(survey)
```
Now that we have all the variables, let's show off what kind of variables they are.
Sex: The sex of a student, with the options of female and male. This is a categorical variable, and it's a binary variable since it's only between two options. 

Nonwriting Hand: This is the span of the non writing hand from tip of the thumb to the tip of the pinkie in centimeters. This is a quantitative variable, and it's continuous since it's measured of a non finite value. 

Writing Hand: This is the span of the writing hand from tip of the thumb to the tip of the pinkie in centimeters. This is a quantitative variable, and it's continuous since it's measured of a non finite value.

Fold: Folding your hands, with the options being right on left, left on right, or neither. This is a categorical variable, and it's a nominal since there is no real rank or order between the groups.

Pulse: This is beats per minute of students. This is a quantitative variable, and it's discrete since this is an integer count of individual values. 

Top hand of clap: When clapping one hand is on top, this records whether the hand is the right, the left, or neither. This is a categorical variable, and it's nominal since there is no real rank or order between the groups.

Exercise: The amount that a student exercises, with the options being frequently, some, and none. This is a categorical variable, and it's an ordinal variable since the groups can be ranked inside of a specific order of least amount of exercise to greatest. 

Smoke: The amount that a student smokes, which varies between never, regular, occasionally, and heavy. This is a categorical variable, and it's ordinal since the groups can be ranked inside of a specific order of least smoking to greatest. 

Height: This is the height of the student in centimeters. It is a quantitative variable, and it is continuous since it measures in non infinite values and it's measurable. 

Metric versus Imperial: This variable represents whether a student represents their height using the metric system or using the imperial system. This is a categorical variable, and it's binary as there's only two options. 

Age: This is the age of the student in years. It is a quantitative variable, and it is continuous since it measures in non infinite values and it's measurable. 
span of writing hands, one compare it with span of non-writing hand and one compare it with height of students.
however, another plot i compared between pulse and age in terms of exercise frequency
The outcome variable for question three for the first graph is writing hand span, the second graph it's pulse rate, and the third graph is hand span. 

# Q2
First, we import that dataset and remove and filter-out any rows with missing variables.
```{r}
survey <- read.csv("survey.csv")
survey[survey == ''] <- NA
survey <- survey %>% drop_na()
```

The first contingency table compares the level of exercise and smoking in students. Majority of people never smoke. Also, even among those that smoke, most exercise sometimes or frequently.
```{r}
table(data.frame("Exer" = survey$Exer, "Smoke" = survey$Smoke))
```

The second contingency table compares writing hand with the upper hand when clapping. Most students are right-handed and most right-handed students clap with their right hand on top.
```{r}
table(data.frame("W.Hnd" = survey$W.Hnd, "Clap" = survey$Clap))
```

The first graph is a boxplot of student heights(Both overall and also divided by sex). There's a median of 171, a max of 200, and a min of 150 for the entire population. Also male students are, on average, taller compared to female students. Also it can be seen that the tallest person is male and shortest person is female.
```{r}
males <- survey %>% filter(Sex == "Male")
females <- survey %>% filter(Sex == "Female")
heights <- data.frame("All Students" = survey$Height,
                      "Male Students" = males$Height,
                      "Female Students" = females$Height)

boxplot(heights, main = "Student Height Boxplot", ylab = "Height(cm)")
```

The second graph is a histogram of student ages. Most people are under 20, vast majority of people are under 25, and the oldest people are over 70.
```{r}
hist(survey$Age,
     breaks=20,
     main = "Student Age Histogram",
     xlab = "Age")
```

Finally, you can see the summary statistics for 2 variables. Student pulse rate(beats per minute) and student writing hand span(distance from tip of thumb to tip of little finger of spread hand).
```{r}
summary(survey$Pulse)
summary(survey$Wr.Hnd)
```

# Q3
## First multivariate visualization
Check how does `Sex` influence the relationship between `Wr.Hnd` and `Height` (the units are all centimeters), where `Wr.Hnd` represents the distance of span from tip of thumb to little finger of student's writing hand.\
By expecting `Height` as the explanatory variable and `Wr.Hnd` as the response variable, we are looking for the relationship if student with higher heights would have larger writing span. By using different color to distinguish between gender: _red_ represents female and _green_ represents male. By having regression line shade could give clear visualization.By drawing the regression lines to have a clear picture of the analysis, without any significant difference between male and female students. Since female students are shorter than male students, and the height of male students are varies more than female students.\
However, both of male and female students are all have the similar positive correlation, where the higher of student's height, the larger their span of writing hands would be.
```{r}
ggplot(survey, aes(x = Height, y = Wr.Hnd)) +
  geom_point(aes(color = Sex), size = 2) +
  geom_smooth(aes(color = Sex), formula = y ~ x, method = "lm") +
  labs(title = "Span of Writing Hands Variers with Heights by Sex",
       x = "Height of Students (cm)", y = "Span of Writing Hands (cm)")
```

## Second multivariate visualization
Based on the data analysis previously, we have noticed that majority of participants' age range from 16 to 21, out main focus here is to see the tendencies of majority people, by filtering out participants who are not in this range would help us see the main focus of study better.\
Here, we check how does `Exer` influence the relationship between `Age` and `Pulse`, where `Exer` represents the frequency that students exercise and `Pulse` represents their pulse beats per minute.
Because we want to looking for the meaningful relationship of whether the increase of age would influence students' pulse rate in terms of how often they exercise. Therefore, we let `Age` as the explanatory variable and `Pulse` as the response variable.\
Unlike previous visualization, we separate our graph into three small group by `Exer` to have better looking into the result. We also use different color to distinguish `None`, `Freq`, and `some` for Exercises: by using _green_ for students who do exercises frequently, _orange_ for students who don't do exercises, and _purple_ for students who sometimes do exercises. With the regression lines, we have noticed the significant difference of `Age` and `Pulse` between students who have never do exercises (none in `Exer`) with students who do exercises (frequently or some in `Exer`). With students who never do exercises, they have positive correlation, while other students who do exercises don't have significant relationships. Equivalently to say, students who don't do exercises, as the age grow, their heart pulse tend to beat faster per minute; and students who do exercises frequently or sometimes, their heart beat to age don't have significant relationship (for students who do **some** exercises have little negative correlation, as the age grows, their heart pulse would decrease, but not significant comparing to students who have **none** exercises.)
```{r}
age_range <- survey %>% filter(between(Age, 16, 21))

plt <- ggplot(age_range, aes(x = Age, y = Pulse)) +
  geom_point(aes(color = Exer), size = 3, color = "#69b3a2", pch = 18) +
  geom_smooth(aes(color = Exer),
              formula = y ~ x, method = "lm",
              se=FALSE, fullrange=TRUE) +
  theme_classic() +
  facet_grid(.~Exer) +
  labs(title = "Pulse Rate by Age and Frequency of Exercises",
       x = "Age of students (years)", y = "Pulse rate of students (beats per minute)")
plt + scale_color_brewer(palette = "Dark2")
```

## Third multivariate visualization
Here, we want to check the relationship between `Wr.Hnd` and `NW.Hnd` varies between `Sex` - female and male students at the University of Adelaide, which `Wr.Hnd` represents the span of student's writing hand in centimeters and `NW.Hnd` represents the span of student's non-writing hand in centimeters.\
The graph below contains different significant data, each dot represents each student's data, and they are colored in _green_ and _orange_ color to distinguish between gender (with orange dots represent male students and green dots represent female students).\ 
By considering `Wr.Hnd` as the explanatory variable and `NW.Hnd` as the response variable because we are looking for how the span size of biological (right-hand or left-hand) hands correlates to the span of non-writing hands.\
Now, by looking at our scatterplot, with the strong positive correlation between span of writing hands and non-writing hands, we can observe that the larger span of student's writing hand, the larger span of student's non-writing hand would become. By having overlapping points between 16.5cm to 19.5cm for writing hands' span, it means majority of students (both female and male students) lie in the range. With some female students have significant smaller span of writing and non-writing hands as the outliers.\
There are significant cluster for male participants, all female students' span of their writing and non-writing hands are below 21cm, with the significant cluster of male students' data lie above 21cm. We can say that male students are generally have larger span of writing and non-writing hands comparing to female students.
```{r}
graph <- ggplot(survey, aes(x = Wr.Hnd, y = NW.Hnd, color = Sex)) +
  geom_point() +
  geom_labelsmooth(aes(label = Sex), fill = "pink", 
                   method = "lm", size = 2, linewidth = 1, 
                   boxlinewidth = 0.3, formula = y ~ x) +
  labs(title = "Span of Hands Comparison by Writing Hands and Sex",
        x = "Span of Writing Hands (cm)", y = "Span of Non-writing Hands (cm)")
graph + scale_color_brewer(palette = "Dark2")
```
