---
title: "STAT-180 Project-1"
author: "Sophia Zhang, Mojdeh Motalebi Kashani"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```

# Q2
First, we import that dataset and remove and filter-out any rows with missing variables.
```{r}
survey <- read.csv("survey.csv")
survey[survey == ''] <- NA
survey <- survey %>% drop_na()
```

The first contingency table compares the level of exercise and smoking in students. Majority of people never smoke. Also, even among those that smoke, most exercise sometimes or frequently.
```{r}
table(data.frame("Exer" = survey$Exer, "Smoke" = survey$Smoke))
```

The second contingency table compares writing hand with the upper hand when clapping. Most students are right-handed and most right-handed students clap with their right hand on top.
```{r}
table(data.frame("W.Hnd" = survey$W.Hnd, "Clap" = survey$Clap))
```

The first graph is a boxplot of student heights. With a median of 171, a max of 200, and a min of 150.  
```{r}
boxplot(survey$Height, main = "Student Height")
```

The second graph is a histogram of student ages. Most people are under 20, vast majorty of people are under 25, and the oldest people are over 70.
```{r}
hist(survey$Age, main = "Student Age")
```

Finally, you can see the summary statistics for 2 variables. Student pulse rate(beats per minute) and student writing hand span(distance from tip of thumb to tip of little finger of spread hand).
```{r}
summary(survey$Pulse)
summary(survey$Wr.Hnd)

# Q4
## First multivariate visualization
Check how does `Sex` influence the relationship between `Wr.Hnd` and `Height` (the units are all centimeters), where `Wr.Hnd` represents the distance of span from tip of thumb to little finger of student's writing hand.
By expecting `height` as the explanatory variable and `Wr.Hnd` as the response variable, we are looking for the relationship if student with higher heights would have larger writing span.
By drawing the regression lines to have a clear picture of the analysis, without any significant difference between male and female students. However, both of male and female students are all have the similar positive scatterplotss, where the higher of student's height, the larger their span of writing hands would be.
```{r}
ggplot(survey, aes(x = Height, y = Wr.Hnd)) +
  geom_point(aes(color = Sex), size = 2) +
  geom_smooth(aes(color = Sex), method = "lm") +
  labs(title = "Span of Writing hands variers heights by sex of students",
       x = "Height of student (cm)", y = "Span of writing hand (cm)")
```

```
